State of the Practice for Lattice Boltzmann Method Software

### 2022-04-23 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
reviewed Section 8

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
	Development Time} A developer of pyLBM noted that their small development
	team has a lack of time to implement new features. Small development teams
	are common for LBM software packages (as shown in the measurement table
	excerpt in Figure~\ref{measurement_template_image}). Lack of time is also
	highlighted as a pain point by other research software developers
	\citet{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackSoftDevExp}:] \textbf{Lack
	of Software Development Experience} A lack of software development
	experience was noted by the developer of TCLB, and others noted a need for
	improving software engineering education. Many of the team members on their
	project are domain experts, not computer scientists or software engineers.
	This same trend is noted by \citep{Nguyen-HoanEtAl2010}, which showed only
	23\% of research software survey respondents having a computing-related
	background. Similarly, \citep{UditAndKatz2017} show that the majority (54\%)
	of postdocs have not received training in software development.  The LBM
	developer suggesting an increasing role for formal software education
	matches the trend observed by \citet{PintoEtAl2018}, where their replication
	of a previous study \citep{HannayEtAl2009}, shows a growing interest in
	formal training (From 13\% of respondents in 2009 to 22\% in 2018).
	\citep{PintoEtAl2018} found that some developers feel there is a mismatch
	between coding skills and subject-matter skills.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
	Incentive and Funding} The TCLB developer noted a lack of incentives and
	funding in academia for developing widely used research software. This
	problem has also been noted by others \citep{gewaltig2012quality, Goble2014,
	KaterbowAndFeulner2018}.  \citet{WieseEtAl2019} reported developer pains
	related to publicity, since historically publishing norms make it difficult
	to get credit for creating software.  As studied by
	\citet{HowisonAndBullard2016}, research software (specifically biology
	software, but the trend likely applies to other research software domains)
	is infrequently cited.  \citep{PintoEtAl2018} also mentions the lack of
	formal reward system for research software.

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
	\textbf{Technology Hurdles} Technology pain points for LBM developers
	include setting up parallelization and CI. The pain
	point survey of \citep{WieseEtAl2019} also highlighted technical-related
	problems like dependency management, cross-platform compatibility,
	CI, hardware issues and operating system issues.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
	\textbf{Ensuring Correctness} Difficulties with ensuring correctness were
	noted by several developers. They alluded to difficulty with testing the
	correctness of large numbers of features, and challenges with automated
	testing. The TCLB developer commented that the amount of testing
	data needed was sometimes problematic, since free testing services do not
	offer adequate facilities for large amounts of data, which means in-house
	testing solutions are needed.  Other research software domains point to the
	following problems with testing: i) \citep{PintoEtAl2018} mention the
	problem of insufficient testing; ii) the survey of \citep{HannayEtAl2009}
	shows that more developers think testing is important than the number that
	believe they have a sufficient understanding of testing concepts; and, iii)
	\citep{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}
	point to the oracle problem, which occurs in research software when we do
	not have a means to judge the correctness of the calculated solutions.  The
	LBM experience seems to overlap with i and ii, but the LBM developers did
	not allude to problem iii (the oracle problem) in our conversations.  As
	discussed in Section~\ref{SecReproducibility}, the oracle problem likely did
	not come up because the LBM developers we interviewed have strategies for
	verifying their work.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:]
	\textbf{Usability} Several developers noted that users sometimes try to use
	incorrect LBM method combinations to solve their problems. Furthermore, some
	users think that the packages will work out of the box to solve their cases,
	while in reality CFD knowledge needs to be applied to correctly modify the
	packages for the new endeavour.  Some respondents in the survey of
	\citep{WieseEtAl2019} also mentioned that users do not always have the
	expertise required to install or use the software.

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechDebt}:] \textbf{Technical
	Debt} The developer of ESPResSo said that their source code was written with
	a specific application in mind, which later caused too much coupling between
	components in the source code. This results in technical debt
	\citet{KruchtenEtAl2012}, which has an impact on future maintainability and
	reusability. Concern with technical debt is likely why researchers in the
	survey of \citep{Nguyen-HoanEtAl2010} rated maintainability as the third
	most important software quality. More recently the push for sustainable
	software \citet{deSouzaEtAl2019} is motivated by the pain that past
	developers have had with accumulating too much technical debt.

>>>>

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
	Development Time} A developer of pyLBM noted that their small development
	team has a lack of time to implement new features. Small development teams
	are common for LBM software packages (as shown in the measurement table
	excerpt in Figure~\ref{measurement_template_image}). Lack of time is also
	highlighted as a pain point by other research software developers
	\citep{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackSoftDevExp}:] \textbf{Lack
	of Software Development Experience} A lack of software development
	experience was noted by the developer of TCLB, and others noted a need for
	improving software engineering education. Many of the team members on their
	project are domain experts, not computer scientists or software engineers.
	This same trend is noted by \citet{Nguyen-HoanEtAl2010}, which showed only
	23\% of research software survey respondents having a computing-related
	background. Similarly, \citet{UditAndKatz2017} show that the majority (54\%)
	of postdocs have not received training in software development.  The LBM
	developer suggesting an increasing role for formal software education
	matches the trend observed by \citet{PintoEtAl2018}, where their replication
	of a previous study \citep{HannayEtAl2009}, shows a growing interest in
	formal training (From 13\% of respondents in 2009 to 22\% in 2018).
	\citet{PintoEtAl2018} found that some developers feel there is a mismatch
	between coding skills and subject-matter skills.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
	Incentive and Funding} The TCLB developer noted a lack of incentives and
	funding in academia for developing widely used research software. This
	problem has also been noted by others \citep{gewaltig2012quality, Goble2014,
	KaterbowAndFeulner2018}.  \citet{WieseEtAl2019} reported developer pains
	related to publicity, since historically publishing norms make it difficult
	to get credit for creating software.  As studied by
	\citet{HowisonAndBullard2016}, research software (specifically biology
	software, but the trend likely applies to other research software domains)
	is infrequently cited.  \citet{PintoEtAl2018} also mentions the lack of
	formal reward system for research software.

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
	\textbf{Technology Hurdles} Technology pain points for LBM developers
	include setting up parallelization and CI. The pain
	point survey of \citet{WieseEtAl2019} also highlighted technical-related
	problems like dependency management, cross-platform compatibility,
	CI, hardware issues and operating system issues.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
	\textbf{Ensuring Correctness} Difficulties with ensuring correctness were
	noted by several developers. They alluded to difficulty with testing the
	correctness of large numbers of features, and challenges with automated
	testing. The TCLB developer commented that the amount of testing
	data needed was sometimes problematic, since free testing services do not
	offer adequate facilities for large amounts of data, which means in-house
	testing solutions are needed.  Other research software domains point to the
	following problems with testing: i) \citep{PintoEtAl2018} mention the
	problem of insufficient testing; ii) the survey of \citep{HannayEtAl2009}
	shows that more developers think testing is important than the number that
	believe they have a sufficient understanding of testing concepts; and, iii) previous studies
	\citep{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}
	point to the oracle problem, which occurs in research software when we do
	not have a means to judge the correctness of the calculated solutions.  The
	LBM experience seems to overlap with i and ii, but the LBM developers did
	not allude to problem iii (the oracle problem) in our conversations.  As
	discussed in Section~\ref{SecReproducibility}, the oracle problem likely did
	not come up because the LBM developers we interviewed have strategies for
	verifying their work.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:]
	\textbf{Usability} Several developers noted that users sometimes try to use
	incorrect LBM method combinations to solve their problems. Furthermore, some
	users think that the packages will work out of the box to solve their cases,
	while in reality CFD knowledge needs to be applied to correctly modify the
	packages for the new endeavour.  Some respondents in the survey of
	\citet{WieseEtAl2019} also mentioned that users do not always have the
	expertise required to install or use the software.

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechDebt}:] \textbf{Technical
	Debt} The developer of ESPResSo said that their source code was written with
	a specific application in mind, which later caused too much coupling between
	components in the source code. This results in technical debt
	\citep{KruchtenEtAl2012}, which has an impact on future maintainability and
	reusability. Concern with technical debt is likely why researchers in the
	survey of \citet{Nguyen-HoanEtAl2010} rated maintainability as the third
	most important software quality. More recently the push for sustainable
	software \citep{deSouzaEtAl2019} is motivated by the pain that past
	developers have had with accumulating too much technical debt.
